---
title: "Darwin Core Mapping"
subtitle: "VMM - Macroinvertebrate occurrences collected by VMM in Flanders, Belgium"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(rgbif)          # To use GBIF services
```

# Read source data

```{r}
input_event <- read_excel(here::here("datasets", "vmm-macroinvertebrates-events", "data", "interim", "210112FL_RV_events.xlsx"))

input_occurrence <- read_excel(here::here("datasets", "vmm-macroinvertebrates-events", "data", "interim", "210112FL_RV_occurrence.xlsx"))

input_meas <- read_excel(here::here("datasets", "vmm-macroinvertebrates-events", "data", "interim", "210112FL_RV_MeasurementOrFact.xlsx"))
```

Preview data:

```{r}
input_event %>% head()
input_meas %>% head()
input_occurrence %>% head()
```

# Preprocessing

## Tidy data

Clean data somewhat:

```{r}
input_event %<>% 
  remove_empty("rows") %<>%
  clean_names()

input_meas %<>% 
  remove_empty("rows") %<>%
  clean_names()

input_occurrence %<>% 
  remove_empty("rows") %<>%
  clean_names()
```


# Darwin Core mapping

# occurrence

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

```{r}
head(input_occurrence, n =10)
```


```{r}
occurrence <- input_occurrence
```


### eventID

```{r}
occurrence %<>% mutate(dwc_eventID = event_id)
```


### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### occurrenceID


```{r}
occurrence %<>%  mutate(dwc_occurrenceID = occurrence_id)
```

### recordedBy

**CHECK IN EMAILED FILE**

```{r}
#occurrence %<>% mutate(dwc_recordedBy = recorded_by)
```

### organismQuantity

```{r}
occurrence %<>% mutate(dwc_organismQuantity = individual_count)
```

### organismQuantityType

```{r}
occurrence %<>% mutate(dwc_organismQuantityType = "countRange")

```

### scientificName

```{r}
occurrence %<>% mutate(dwc_scientificName = scientific_name)
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```

### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = taxon_rank)
```

### nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN")
```

### taxonRemarks

```{r}
occurrence %<>% mutate(dwc_taxonRemarks = taxon_remarks)
```

# Event Core

### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en")
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/")
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "VMM")
```

### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "insert doi")
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "VMM")
```

### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "MICA - Muskrat occurrences collected by VMM in Flanders, Belgium")
```






### individualCount

```{r}
occurrence %<>% rename(dwc_individualCount = vangst_aantal) 
```

### occurrenceStatus

```{r}
occurrence %<>% mutate(dwc_occurrenceStatus = case_when(
  dwc_individualCount > 0 ~ "present",
  dwc_individualCount == 0 ~ "absent"
))
```

### eventID

```{r}
occurrence %<>% mutate(dwc_eventID = str_c("MICA:VMM:EV:", registratie_id))
```

### eventDate

```{r}
occurrence %<>% mutate(dwc_eventDate = as.character(dag))
```

### samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingProtocol = "rat trap") 
```

### locationID

```{r}
occurrence %<>% mutate(dwc_locationID = locatie_id)
```

### continent

```{r}
occurrence %<>% mutate(dwc_continent = "Europe")
```

### countryCode

```{r}
occurrence %<>% mutate(dwc_countryCode = "BE")
```

### stateProvince

```{r}
occurrence %<>% mutate(dwc_stateProvince = recode(provincie_omschrijving,
  "Antwerpen" = "Antwerp",
  "Limburg" = "Limburg",
  "Oost-Vlaanderen" = "East Flanders",
  "Vlaams-Brabant" = "Flemish Brabant",
  "West-Vlaanderen" = "West Flanders",
  "Onbekend" = NA_character_
))
```

### municipality

```{r}
occurrence %<>% mutate(dwc_municipality = recode(gemeente_naam,
  "Onbekend" = NA_character_
))
```

### locationRemarks

```{r}
occurrence %<>% mutate(dwc_locationRemarks = recode(vha_categorie_omschrijving, 
  "CAT2 - Onbevaarbaar cat. 2" = "CAT2 - not navigable cat. 2",
  "CAT1 - Onbevaarbaar cat. 1" = "CAT1 - not navigable cat. 1",
  "CAT3 - Onbevaarbaar cat. 3" = "CAT3 - not navigable cat. 3",
  "CAT - Andere" = "CAT - other",
  "ONB - Onbekend" = "ONB - unknown",
  "BEV - Bevaarbaar" = "BEV - navigable"
))
```

### decimalLatitude

```{r}
occurrence %<>% mutate(dwc_decimalLatitude = sprintf(
  "%.5f",
  round(locatie_gps_breedte, digits = 5)
))
```

### decimalLongitude

```{r}
occurrence %<>% mutate(dwc_decimalLongitude = sprintf(
  "%.5f",
  round(locatie_gps_lengte, digits = 5)
))
```

### geodeticDatum

```{r}
occurrence %<>% mutate(dwc_geodeticDatum = "WGS84") 
```

### coordinateUncertaintyInMeters

```{r}
occurrence %<>% mutate(dwc_coordinateUncertaintyInMeters = "30")
```





### vernacularName

```{r}
occurrence %<>% mutate(dwc_vernacularName = vernacularName)
```




## Post-processing

```{r}
occurrence %<>% select(starts_with("dwc_"))
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "mica-vmm-occurrences", "data", "processed", "occurrence.csv"), na = "")
```
