---
title: "Darwin Core Mapping"
subtitle: "Eikenprocessierups in East-Flanders, Belgium"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # for the date
```

# Read source data

Create a data frame `input_data` from the source data:
The source data was provided in Excel


```{r}
input_interimExcel <- read_excel(here::here("datasets", "eikenprocessierups-occurrences", "data", "raw", "2022_gegevens_EPR_1.xlsx" ))

input_interimRef <- read_delim(here::here("datasets", "eikenprocessierups-occurrences", "data", "interim", "EPRgegevensRAWRefine.csv" ),',')

input_interim <- read_delim(here::here("datasets", "eikenprocessierups-occurrences", "data", "interim", "2022_12_14EPRgegevensRAWRefine.csv" ),',')


```

Preview data:

```{r}
input_interim %>% head(n = 5)
```

# Process source data

## Tidy data

Clean data somewhat:


```{r}
input_interim <-
  here::here(
    "datasets",
    "eikenprocessierups-occurrences",
    "data",
    "interim",
    "2022_12_12EPRgegevensRAWRefine.csv"
  ) %>% readr::read_csv() %>%
  mutate(across(everything(),  ~ stringr::str_replace(.x, "\n", " ")))
```



```{r}
input_interim %<>% remove_empty("rows")
```

```{r}
input_interim %>% head(n = 5)
```

# Georeference source data

```{r}
input_interimDec <- input_interim 
```

# keep only decimal data lat long

```{r}
input_interimDec %<>% mutate(verbatimLongitude = X) %>%
                   mutate(verbatimLatitude = Y) 
                   
```

```{r}
input_interimDec  %<>% filter(!is.na(decimallongitude))
```



## Lambert to Decimals

### Keep original x,Y as verbatimcoordinates

We want to keep the original verbatim Coordinates
The original X,Y will be used for creating decimalLatitude and decimalLongitude

```{r}
input_interim %<>% mutate(verbatimLongitude = X) %>%
                   mutate(verbatimLatitude = Y) 
                   
```

### Create a spacial dataframe

Create spatial dataframe and define the CRS (31370 is Belgian Lambert 72)
We indicate the x,y columns as coordinates and define the coordinate system

```{r}
input_interim  %<>% filter(!is.na(X))
```


```{r}
input_interim = st_as_sf(input_interim, coords = c("X","Y"), crs = 31370 , agr = "constant", na.fail = TRUE )
input_interim %>% head(n = 5)
```

## Transform the data in WGS84

Now we transform the Lambert 72 in WGS84

```{r}
input_interim <- st_transform(input_interim, crs = 4326)
input_interim %>% head(n = 5)
```

### Create a dataframe with column decimalLatitude & decimalLongitude

In order to add the decimal coordinates into the original df we create an x_y dataframe

```{r}
x_y <- as.data.frame(st_coordinates(input_interim))
x_y %>% head(n = 5)

write_csv(x_y, here::here("datasets", "eikenprocessierups-occurrences", "data", "interim", "x_y.csv"), na = "")
```

# Create extra dataframe to work with 

```{r}
input_interim_df <- input_interim
```

# Remove geometry from dataframe 'input_interim_df'

We remove all geometry from the spacial dataframe, to create a normal df

```{r}
st_geometry(input_interim_df) <- NULL

input_interim_df %>% head(n = 5)
```

### Bind columns x_y and input_interim_df

Join both df's to have the decimalLat & Lon

```{r}
input_interim_df = bind_cols(input_interim_df,x_y, .id = NULL)

input_interim_df %>% head(n = 5)
```

### Integrate x-y coordinates in original dataframe

### collate the dec-lat rows



We can safely substitute input_interim 

```{r}
input_interim <- input_interim_df
```



reinstitute the missing rows

```{r}
head(input_interimDec)

head(input_interim)
```

```{r}
input_interim %<>%
  mutate(X = round(X, digits = 5)) %>%
  mutate(Y = round(Y, digits = 5))

```
```{r}
#input_interim %<>% select(-c( X, Y))
```


```{r}
input_interim <- mutate_all(input_interim, as.character)
```


```{r}
head(input_interimDec)

head(input_interim)
```



```{r}
input_interim <- union(input_interimDec, input_interim)
```


```{r}
input_interim %<>% mutate(dwc_decimalLatitude2 = coalesce(Y, decimalLatitude ))
```
```{r}
input_interim %<>% mutate(dwc_decimalLongitude2 = coalesce(X, decimallongitude ))
```

# Occurrence core

## Pre-processing

Create a dataframe occurrence data only 

```{r}
occurrence <- input_interim
```

## remove obsolete columns

```{r}
occurrence %<>% select(-c(ADRES, decimallongitude, decimalLatitude, adres2, Melder, X, Y))
```


# Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).


# Event

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "to complete")
```

### add row numbers to genrate occurrenceID unique


```{r}
occurrence %<>% mutate(ID_row = row_number())
```


### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "Province East-Flanders") # e.g. "INBO"
```
### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
#occurrence %<>% mutate(datasetID = "insert doi") 
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "POV") # e.g. "INBO"
```

### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Monitoring of Oak processionary (Thaumetopoea processionea) in the province of East Flanders, Belgium") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

The following terms contain information about the taxon:

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### informationWithHeld

### dataGeneralizations

### occurrenceID

```{r}
occurrence %<>% rename(dwc_occurrenceID = 'Nr') %>%
                mutate(dwc_occurrenceID = str_c("POV:EPR:", dwc_occurrenceID))
  
```

### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = 'POV')
```

### individualCount


### organismQuantity


```{r}
occurrence %<>%  mutate(dwc_organismQuantity = gemiddeldAantalNestenPerBoom) 
                
```



### organismQuentityType

```{r}
occurrence %<>%  mutate(dwc_organismQuantityType = 'number of nests in a tree') 
                
```

### sex

### lifeStage

### behavior

### occurrenceRemarks

```{r}
#occurrence %<>% mutate(dwc_occurrenceRemarks = str_c('number of infected trees: ', aantalBomen ))
```

## samplingProtocol





### samplingEffort

### year

```{r}
occurrence %<>% mutate(dwc_year = year)
```


### eventDate

eventDate already ok in source file (should be)

```{r}

occurrence %<>% mutate(dwc_eventDate = eventDate)

#occurrence %<>% mutate(dwc_eventDate = as_date(ymd(eventDate) , format = "%Y-%m-%d")) 

#occurrence %<>% mutate(dwc_eventDate2 = str_replace_all(dwc_eventDate, " ", "0"))

#occurrence %<>% ymd(date)

#occurrence %<>% mutate(dwc_eventDate = dmy(eventDate))
# mutate(eventDate = as_date(ymd(date))) # , format = "%Y-%m-%d")) 

head(occurrence, n = 5)
```

# verbatimEventDate

```{r}
occurrence %<>% mutate(dwc_verbatimEventDate = verbatimEventDate )
```


# Location

```{r}
occurrence %<>%
  mutate(dwc_decimalLongitude = dwc_decimalLongitude2) %>%
  mutate(dwc_decimalLatitude = dwc_decimalLatitude2) %>%
  mutate(dwc_geodeticDatum = "WGS84") %>%
  mutate(dwc_coordinateUncertaintyInMeters = "30") %>%
  mutate(dwc_verbatimCoordinateSystem = "Lambert coordinates") %>%
  mutate(dwc_verbatimSRS = "Belgian Datum 1972") %>%
  mutate(dwc_countryCode = "BE")  %>%            
  mutate(dwc_continent = "Europe")
```

```{r eval=FALSE, include=FALSE}
head(occurrence, n = 5)
occurrence %<>%
  mutate(dwc_verbatimLongitude = round(verbatimLongitude)) %>%
  mutate(dwc_verbatimLatitude = round(verbatimLatitude)) %>%
  mutate(dwc_decimalLongitude = round(dwc_decimalLongitude, digits = 5)) %>%
  mutate(dwc_decimalLatitude = round(dwc_decimalLatitude, digits = 5))
```
```{r}
occurrence %<>%   
   mutate(dwc_decimalLatitude = as.character(format(dwc_decimalLatitude, nsmall = 5))) %>%
   mutate(dwc_decimalLongitude = as.character(format(dwc_decimalLongitude, nsmall = 5)))
```

### continent

```{r}
##occurrence %<>% mutate(continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
#occurrence %<>% mutate(countryCode = "BE") # e.g. "Belgium = BE"
```

### municipality

```{r}
occurrence %<>% mutate(dwc_municipality = Gemeente) # e.g. "Belgium = BE"
```


### verbatimcoordinates

### verbatimLatitude

### verbatimLongitude

### verbatimcoordinatesystem

```{r}
#occurrence %<>% mutate(dwc_verbatimcoordinatesystem = "Lambert coordinates") # 
```

### verbatimSRS

```{r}
#occurrence %<>% mutate(verbatimSRS = "Belgian Datum 1972")
```

### decimalLatitude

### decimalLongitude

### geodeticDatum

### coordinateUncertaintyInMeters

### georeferenceRemarks

### identifiedBy

```{r}
occurrence %<>% mutate(dwc_identifiedBy = 'Province East Flanders')
```

### scientificName 

```{r}
occurrence %<>% mutate(dwc_scientificName = 'Thaumetopoea processionea')
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```

### scientificNameAuthorship

```{r}
occurrence %<>% mutate(dwc_scientificNameAuthorship = 'Linnaeus, 1758' )


```

### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = "species")
```

### nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN") # e.g. "ICZN"
```

### occurrenceStatus

```{r}
occurrence %<>% 
    ##select(individualCount) %>%
    mutate(dwc_occurrenceStatus = 'present')

```




## Post-processing

```{r}

occurrence %<>% select(starts_with("dwc_"))
```



```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")

```

## remove obsolete columns

```{r}
occurrence %<>% select(-c(decimalLatitude2, decimalLongitude2))
```

```{r}
occurrence <- mutate_all(occurrence, as.character())
```



Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_delim(occurrence, here::here("datasets", "eikenprocessierups-occurrences", "data", "processed", "occurrence.csv"), na = "", delim = ',')
```


