---
title: "Darwin Core Mapping"
subtitle: "Eikenprocessierups in East Flanders, Belgium"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # for the date
```

# Read source data

Create a data frame `input_data` from the source data:
The source data was provided in Excel


```{r}
input_interimExcel <- read_excel(here::here("datasets", "eikenprocessierups-occurrences", "data", "raw", "2022_gegevens_EPR_1.xlsx" ))

input_interim <- read_delim(here::here("datasets", "eikenprocessierups-occurrences", "data", "raw", "2022_gegevens_EPR_1_a.csv" ),';')

```

Preview data:

```{r}
input_interim %>% head(n = 5)
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_interim %<>% remove_empty("rows")
```

```{r}
input_interim %>% head(n = 5)
```

# Georeference source data


## Lambert to Decimals

### Keep original x,Y as verbatimcoordinates

We want to keep the original verbatim Coordinates
The original X,Y will be used for creating decimalLatitude and decimalLongitude

```{r}
input_interim %<>% mutate(verbatimLongitude = X) %>%
                   mutate(verbatimLatitude = Y) 
                   
```

### Create a spacial dataframe

Create spatial dataframe and define the CRS (31370 is Belgian Lambert 72)
We indicate the x,y columns as coordinates and define the coordinate system

```{r}
input_interim = st_as_sf(input_interim, coords = c("X","Y"), crs = 31370 , agr = "constant" )
input_interim %>% head(n = 5)
```

## Transform the data in WGS84

Now we transform the Lambert 72 in WGS84

```{r}
input_interim <- st_transform(input_interim, crs = 4326)
input_interim %>% head(n = 5)
```

### Create a dataframe with column decimalLatitude & decimalLongitude

In order to add the decimal coordinates into the original df we create an x_y dataframe

```{r}
x_y <- as.data.frame(st_coordinates(input_interim))
x_y %>% head(n = 5)

write_csv(x_y, here::here("datasets", "eikenprocessierups-occurrences", "data", "interim", "x_y.csv"), na = "")
```

# Create extra dataframe to work with 

```{r}
input_interim_df <- input_interim
```

# Remove geometry from dataframe 'input_interim_df'

We remove all geometry from the spacial dataframe, to create a normal df

```{r}
st_geometry(input_interim_df) <- NULL

input_interim_df %>% head(n = 5)
```

### Bind columns x_y and input_interim_df

Join both df's to have the decimalLat & Lon

```{r}
input_interim_df = bind_cols(input_interim_df,x_y, .id = NULL)

input_interim_df %>% head(n = 5)
```

### Integrate x-y coordinates in original dataframe

We can safely substitute input_interim 

```{r}
input_interim <- input_interim_df
```

# Occurrence core

## Pre-processing

Create a dataframe occurrence data only 

```{r}
occurrence <- input_interim
```

## remove obsolete columns

```{r}
occurrence %<>% select(-c(domain_fr, domain_en, domain_nl, kind_fr, kind_en))
```
## filter for muskrat & bruine rat

```{r}
occurrence %<>%
        filter(kind_nl == 'Bruine rat' | kind_nl == 'Muskusrat') 
        
head (occurrence, n = 5 )

```



# Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

# Event

### datasetID

```{r}
occurrence %<>% mutate(datasetID = "https://doi.org/10.15468/5fps96")
```

### add row numbers to genrate occurrenceID unique


```{r}
occurrence %<>% mutate(ID_row = row_number())
```


### type

```{r}
occurrence %<>% mutate(type = "Event")
```

### language

```{r}
occurrence %<>% mutate(language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(rightsHolder = "RATO") # e.g. "INBO"
```
### accessRights

```{r}
occurrence %<>% mutate(accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
#occurrence %<>% mutate(datasetID = "insert doi") 
```

### institutionCode

```{r}
occurrence %<>% mutate(institutionCode = "INBO") # e.g. "INBO"
```

### datasetName

```{r}
occurrence %<>% mutate(datasetName = "MICA - Muskrat occurrences collected by RATO in East Flanders, Belgium") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

The following terms contain information about the taxon:

### basisOfRecord

```{r}
occurrence %<>% mutate(basisOfRecord = "HumanObservation")
```

### informationWithHeld

### dataGeneralizations

### occurrenceID

```{r}
occurrence %<>% rename(occurrenceID = 'id') %>%
                mutate(occurrenceID = str_c("MICA:RATO:", occurrenceID, ":", ID_row))
  
```

### recordedBy

```{r}
occurrence %<>% mutate(recordedBy = 'RATO')
```

### individualCount

```{r}
occurrence %<>%  rename(individualCount= 'action_amount') 
                
```
```{r}


#occurrence$individualCount %<>% replace_na(0)
```

### organismQuantity


```{r}
occurrence %<>%  mutate(organismQuantity= individualCount) 
                
```



### organismQuentityType

```{r}
occurrence %<>%  mutate(organismQuantityType= action_en) 
                
```


### sex

### lifeStage

### behavior

### occurrenceRemarks

```{r}
occurrence %<>% rename(occurrenceRemarks = 'action_en')
```

## samplingProtocol

```{r}
occurrence %<>% mutate(samplingEffort = materials_en)
```

```{r}
occurrence %<>% rename(samplingProtocol = 'materials_en') %>%
 

# extract follow ups to new column
mutate(follow_up = str_extract(samplingProtocol, "Follow-up=[0-9]*")) %>%
# remove follow ups from samplingProtocol
mutate(samplingProtocol = str_remove(samplingProtocol, "Follow-up=[0-9]*")) %>%
# remove leading ", " if present
mutate(samplingProtocol = str_remove(samplingProtocol, "^,\\s")) %>%
# remove trailing ", " if present
mutate(samplingProtocol = str_remove(samplingProtocol, ",\\s$")) %>%
# replace ", , " in ","
mutate(samplingProtocol = str_replace(samplingProtocol,", , ", ", ")) %>%  
# rename follow up
rename(samplingEffort2 = "follow_up") %>% 
# replace follow-up with times visited
mutate(samplingProtocol = str_remove_all(samplingProtocol, "=[0-9]*"))
```

### samplingEffort

```{r}
head(occurrence, n = 5)

occurrence %<>%
  # extract follow ups to new column
  mutate(samplingEffort = str_replace(samplingEffort, "Follow-up=*", "Times-visited=")) %>%
  mutate(samplingEffort = str_replace_all(samplingEffort, "=",":")) %>%
  mutate(samplingEffort = str_replace_all(samplingEffort, ","," |"))
```

### eventDate

eventDate already ok in source file (should be)

```{r}


#occurrence %<>% ymd(date)

occurrence %<>% mutate(eventDate = dmy(date))
# mutate(eventDate = as_date(ymd(date))) # , format = "%Y-%m-%d")) 

head(occurrence, n = 5)
```

# Location

```{r}
occurrence %<>%
  rename(decimalLongitude = X) %>%
  rename(decimalLatitude = Y) %>%
  mutate(geodeticDatum = "WGS84") %>%
  mutate(coordinateUncertaintyInMeters = "30") %>%
  mutate(verbatimCoordinateSystem = "Lambert coordinates") %>%
  mutate(verbatimSRS = "Belgian Datum 1972") %>%
  mutate(countryCode = "BE")  %>%            
  mutate(continent = "Europe")
```

```{r}
head(occurrence, n = 5)
occurrence %<>%
  mutate(verbatimLongitude = round(verbatimLongitude)) %>%
  mutate(verbatimLatitude = round(verbatimLatitude)) %>%
  mutate(decimalLongitude = round(decimalLongitude, digits = 5)) %>%
  mutate(decimalLatitude = round(decimalLatitude, digits = 5))
```

```{r}
occurrence %<>%   
   mutate(decimalLatitude = as.character(format(decimalLatitude, nsmall = 5))) %>%
   mutate(decimalLongitude = as.character(format(decimalLongitude, nsmall = 5)))
```

### continent

```{r}
##occurrence %<>% mutate(continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
#occurrence %<>% mutate(countryCode = "BE") # e.g. "Belgium = BE"
```

### municipality

municipality already in source file

```{r}
occurrence %<>%
  mutate(municipality = str_remove(municipality,"\\(BE\\)"))
```

### verbatimcoordinates

### verbatimLatitude

### verbatimLongitude

### verbatimcoordinatesystem

```{r}
#occurrence %<>% mutate(verbatimcoordinatesystem = "Lambert coordinates") # 
```

### verbatimSRS

```{r}
#occurrence %<>% mutate(verbatimSRS = "Belgian Datum 1972")
```

### decimalLatitude

### decimalLongitude

### geodeticDatum

### coordinateUncertaintyInMeters

### georeferenceRemarks

### identifiedBy

```{r}
occurrence %<>% mutate(identifiedBy = 'RATO')
```

### scientificName 

```{r}
occurrence %<>% mutate(scientificName = case_when(kind_nl == 'Muskusrat' ~ "Ondatra zibethicus",
                                   kind_nl == 'Bruine rat' ~ "Rattus norvegicus"
                                       )
                                )
```

### kingdom

```{r}
occurrence %<>% mutate(kingdom = "Animalia")
```

### scientificNameAuthorship

```{r}
occurrence %<>% mutate(scientificNameAuthorship = case_when(kind_nl == 'Muskusrat' ~ "Linnaeus, 1766",
                                   kind_nl == 'Bruine rat' ~ "Berkenhout, 1759"
                                       )
                                )


```

### taxonRank

```{r}
occurrence %<>% mutate(taxonRank = "species")
```

### nomenclaturalCode

```{r}
occurrence %<>% mutate(nomenclaturalCode = "ICZN") # e.g. "ICZN"
```

### occurrenceStatus

```{r}
occurrence %<>% 
    ##select(individualCount) %>%
    mutate(occurrenceStatus = case_when(action_nl == 'Vastgesteld' & individualCount > 0 ~ "Present",
                                        action_nl == 'Vastgesteld' & individualCount == 0 ~ "Absent",
                                        action_nl == 'Gevangen' & individualCount > 0 ~ "Present",
                                        action_nl == 'Gevangen' & individualCount == 0 ~ "Absent",
                                        action_nl == 'Verwijderd' & individualCount > 0 ~ "Present",
                                        action_nl == 'Verwijderd' & individualCount == 0 ~ "Absent",
                                        action_nl == 'Geen vangst' & individualCount > 0 ~ "Present",
                                        action_nl == 'Geen vangst' & individualCount == 0 ~ "Absent",
                                        action_nl == 'Sporen' & individualCount > 0 ~ "Present",
                                        action_nl == 'Sporen' & individualCount == 0 ~ "Absent",
                                        action_nl == 'Hoeveelheid'& individualCount > 0 ~ "Present",
                                        action_nl == 'Hoeveelheid'& individualCount == 0 ~ "Absent",
                                        individualCount > 0 ~ "Present",
                                        individualCount == 0 ~ "Absent",
                                        is.na(action_nl) | is.na(individualCount) ~ "Absent"
                                                  
                                       ))
head(occurrence, n = 5)
```

```{r eval=FALSE, include=FALSE}

occurrence %<>%
        filter(eventDate > '2018-01-12' & occurrenceStatus == "Present" & scientificName == "Ondatra zibethicus") 
        
head (occurrence, n = 5 )

```






## Post-processing

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
##occurrence %<>% select(-c(disposition, etiket_ref, collection, scientificName)) %<>% #remove collection columns
          ##  rename(scientificName = verbatimScientificName) 
```

```{r}
occurrence <- mutate_all(occurrence, as.character())
```

Define the order of the output columns

```{r}
col_order <- c( "type","language","license","rightsHolder","accessRights","datasetID"
               ,"institutionCode","datasetName","basisOfRecord","occurrenceID","recordedBy"
               ,"individualCount","organismQuantity","organismQuantityType","occurrenceStatus","occurrenceRemarks"
               ,"samplingProtocol","samplingEffort"
               ,"eventDate","continent","countryCode","municipality"
               ,"verbatimLatitude","verbatimLongitude","verbatimCoordinateSystem","verbatimSRS"
               ,"decimalLatitude","decimalLongitude","geodeticDatum","coordinateUncertaintyInMeters"
               ,"identifiedBy","scientificName","kingdom","scientificNameAuthorship","taxonRank","nomenclaturalCode"
               )
occurrence <- occurrence[, col_order]
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "mica-rato-occurrences", "data", "processed", "occurrence.csv"), na = "")
```
