---
title: "Darwin Core Mapping"
subtitle: "Mediterranean Medetera"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(rgbif)          # To use GBIF services
library(uuid)           # To generate UUID's
```

# Read source data

```{r}
input_data <- read_delim(here::here("datasets", "mediteranian-medetara-occurrences", "data", "interim", "Medetera_occurrences_complete.txt"), "\t")
```

Preview data:

```{r}
input_data %>% head()
```

# Preprocessing

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% 
  remove_empty("rows") %<>%
  clean_names()
```




### fix unique ID issue

```{r}
input_data <- mutate_all(input_data, as.character())
```


# Darwin Core mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

```{r}
occurrence <- input_data
```

### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en")
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/")
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "Marc Pollet")
```

### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "")
```

### institutionCode


### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Mediteranian Medetera")
```

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### occurrenceID


```{r}
occurrence %<>% mutate(dwc_occurrenceID = str_c("Med:med:" , occurrence_id))
```




### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = collector_nm)
```

### individualCount

```{r}
#mutate(occurrence$mm)as.numeric()
  
occurrence %<>% mutate(male = occurrence$mm <- as.numeric(as.character(occurrence$mm)))  %<>%
                mutate(female = occurrence$mm <- as.numeric(as.character(occurrence$ff)))  


```
```{r}
occurrence %<>% mutate(dwc_individualCount = male + female)
```




### occurrenceStatus

```{r}
occurrence %<>% mutate(dwc_occurrenceStatus = case_when(
  dwc_individualCount > 0 ~ "present",
  dwc_individualCount == 0 ~ "absent"
))
```

### eventID

```{r}
occurrence %<>% mutate(dwc_eventID = str_c("Med:MED:ev", sample_cd))
```

### eventDate

```{r}
occurrence %<>% mutate(dwc_eventDate_begin = as.Date(sampling_date_e))
```
```{r}
occurrence %<>% mutate(dwc_eventDate_end = as.Date(sampling_date_b))
```

### samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingProtocol = recode(method_cd
                                                     ,"SW" = "sweep"
                                                     ,"YPT" = "..."
                                                     ,"MT" = "..."
                                                     ,"RsPT" = "..."
                                                     ,"HC" = "..."
                                                     ,"YPT/RPT" = "..."
                                                     ,"WPT" = "..."
                                                     ,"BPT" = "..."
                                                     )) 
```
### locationID

```{r}
occurrence %<>% mutate(dwc_locationID = sampling_site_cd)
```


### locality

```{r}
occurrence %<>% mutate(dwc_locality = locality_nm)
```

### continent

```{r}
occurrence %<>% mutate(dwc_continent = "Europe")
```

### countryCode

```{r}
occurrence %<>% mutate(dwc_countryCode = country_cd)
```

### stateProvince

```{r}
occurrence %<>% mutate(dwc_stateProvince = province_cd
)
```

### municipality

```{r}
occurrence %<>% mutate(dwc_municipality = municipality_pt_nm)
```

### locationRemarks

```{r}
occurrence %<>% mutate(dwc_locationRemarks = recode(vha_categorie_omschrijving, 
  "CAT2 - Onbevaarbaar cat. 2" = "CAT2 - not navigable cat. 2",
  "CAT1 - Onbevaarbaar cat. 1" = "CAT1 - not navigable cat. 1",
  "CAT3 - Onbevaarbaar cat. 3" = "CAT3 - not navigable cat. 3",
  "CAT - Andere" = "CAT - other",
  "ONB - Onbekend" = "ONB - unknown",
  "BEV - Bevaarbaar" = "BEV - navigable"
))
```

### decimalLatitude

```{r}
occurrence %<>% mutate(dwc_decimalLatitude = sprintf(
  "%.5f",
  round(locatie_gps_breedte, digits = 5)
))
```

### decimalLongitude

```{r}
occurrence %<>% mutate(dwc_decimalLongitude = sprintf(
  "%.5f",
  round(locatie_gps_lengte, digits = 5)
))
```

### geodeticDatum

```{r}
occurrence %<>% mutate(dwc_geodeticDatum = "WGS84") 
```

### coordinateUncertaintyInMeters

```{r}
occurrence %<>% mutate(dwc_coordinateUncertaintyInMeters = "30")
```

### scientificName

```{r}
occurrence %<>% mutate(dwc_scientificName = scientificName)
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```

### vernacularName

```{r}
occurrence %<>% mutate(dwc_vernacularName = vernacularName)
```

### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = "species")
```

### nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN")
```

## Post-processing

```{r}
occurrence %<>% select(starts_with("dwc_"))
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
```

```{r}
occurrence <- mutate_all(occurrence, as.character())
```

```{r}
#occurrence %<>% distinct(occurrenceID2, .keep_all = TRUE)
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "mica-vmm-occurrences", "data", "processed", "occurrence.csv"), na = "")
```

# Checks

```{r}
test <- length(unique(occurrence$eventID))
test2 <- length(unique(occurrence$occurrenceID))

n_occur <- data.frame(table(occurrence$occurrenceID))

```

