---
title: "mycrofyten_exoten"
author: "Dimitri Brosens"
date: "18-2-2021"
output: html_document
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # date transformation
library(googlesheets)         # read gsheets
```

# Read source data

Create a data frame `input_data` from the source data:
The source data was corrected in Excel
Muskrat occurrences opened in openRefine
Obsolete columns removed
some columns renamed to DwC term
File exported to csv

```{r}
#input_interim <- read_csv2(file = here:here("datasets", "*****", "data", "raw", "******.csv"))  

# test file copy Dimi
#input_file <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vQ58vi2qvDmdupEwJxHjk2E0vRPq3MED9OfALGq#u01b8LJa9NchiK01WB3aSa5_99hxRj_ptOUbVw90/pub?output=csv")

#input_file <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vSGAgbcqTbCYgtWpm9#-wMnYdhbz_P4QkxG982Hks6rTeNFPWA-wBDq3OoF4YKqY0A/pub?gid=805046402&single=true&output=csv")

 

input_file <- read_excel(here::here("datasets", "invasive-macrofytendata", "data", "raw", "2020-09-22_exoten_macrofytendatabank_waterlopen.xlsx"))

```

Preview data:

```{r}
input_file %>% head(n = 5)
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_interim <- input_file %<>% remove_empty("rows")
```

```{r}
input_interim %>% head(n = 5)
```

# Georeference source data

## split column Standort (includes decimal Lan & long)
Use separate for this

```{r}
#input_interim %<>% 
#         separate(col = Standort, into = c("decimalLatitude", "decimalLongitude"), sep = ",")
#head(input_interim, n = 5)

```


## Lambert to Decimals

### Keep original x,Y as verbatimcoordinates

We want to keep the original verbatim Coordinates
The original X,Y will be used for creating decimalLatitude and decimalLongitude
For this dataset this is part is obsolete (but fun to do!)

```{r}
input_interim %<>% mutate(verbatimLongitude = X) %>%
                   mutate(verbatimLatitude = Y)
```

### Create a spacial dataframe

Create spatial dataframe and define the CRS (31370 is Belgian Lambert 72)(crs = 28992 is the Dutch system used here)
We indicate the x,y columns as coordinates and define the coordinate system

```{r}
input_interim = st_as_sf(input_interim, coords = c("X","Y"), crs = 31370 , agr = "constant" )
input_interim %>% head(n = 5)
```

## Transform the data in WGS84

Now we transform the Lambert 72 in WGS84

```{r}
input_interim <- st_transform(input_interim, crs = 4326)
input_interim %>% head(n = 5)
```

### Create a dataframe with column decimalLatitude & decimalLongitude

In order to add the decimal coordinates into the original df we create an x_y dataframe

```{r}
x_y <- as.data.frame(st_coordinates(input_interim))
x_y %>% head(n = 5)

#write_csv(x_y, here::here("datasets", "mica-uvw-occurrences", "data", "interim/x_y.csv"), na = "")
```

# Create extra dataframe to work with 

```{r}
input_interim_df <- input_interim
```

# Remove geometry from dataframe 'input_interim_df'

We remove all geometry from the spacial dataframe, to create a normal df

```{r}
##input_interim_df <- st_geometry(NULL)

st_geometry(input_interim_df) <- NULL

input_interim_df %>% head(n = 5)
```

### Bind columns x_y and input_interim_df

Join both df's to have the decimalLat & Lon

```{r}
input_interim_df = bind_cols(input_interim_df,x_y, .id = NULL)

input_interim_df %>% head(n = 5)
```

### Integrate x-y coordinates in original dataframe

We can safely substitute input_interim 

```{r}
input_interim <- input_interim_df
```


# use rownumbers as occurrenceID

```{r}
input_interim <-
  input_interim %>%
  rowid_to_column(var = "row_id") %>%
  mutate(rowNumberID = str_c("INBO:MAFY:OCC:", row_id))
```


# Occurrence core

## Pre-processing

Create a dataframe occurrence data only 

```{r}
occurrence <- input_interim
```

## Remove obsolete columns

```{r}
#occurrence %<>% select(-c(kind_nl,day, month, year))
```

# Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

# Event

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "https://doi.orgXXXXXXX") # e.g. "to complete"
```

### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "INBO") # e.g. "INBO"
```

### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```



### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "INBO") # e.g. "INBO"
```

### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "macrofytendatabank_waterlopen") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

The following terms contain information about the taxon:

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### informationWithHeld

### dataGeneralizations

### occurrenceID
```{r}
occurrence %<>% mutate(dwc_occurrenceID = rowNumberID)
```


### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = 'INBO')
```

### individualCount



### organismQuantity

```{r}
occurrence %<>% mutate(dwc_organismQuantity = code) 
```

### organismQuentityType

```{r}
occurrence %<>% mutate(dwc_organismQuantityType = opnameschaal) 
```





### samplingProtocol

```{r}

occurrence %<>% mutate(samplingProtocol = 'inventory')

```


### samplingEffort

### eventDate

```{r}
head(occurrence)

occurrence %<>% mutate(eventDate = datum) 
```

```{r}
occurrence %<>% mutate(year = year(eventDate)) %<>%
                mutate(month = month(eventDate)) %<>%
                mutate(day = day(eventDate)) %<>%
                mutate(eventDate = date(eventDate))
```

# Location

```{r}
head(occurrence)

occurrence %<>% rename(decimalLongitude = X) %>%
                rename(decimalLatitude = Y) %>%
                mutate(geodeticDatum = "WGS84") %>%
                mutate(coordinateUncertaintyInMeters = "30") %>%
                mutate(countryCode = "DE")  %>%            
                mutate(continent = "Europe")
```

```{r}
head(occurrence, n = 5)
occurrence %<>%
#   mutate(verbatimLongitude = round(verbatimLongitude, digits = 0 )) %>%
#   mutate(verbatimLatitude = round(verbatimLatitude, digits = 0)) %>%
   mutate(decimalLongitude = round(decimalLongitude, digits = 5)) %>%
   mutate(decimalLatitude = round(decimalLatitude, digits = 5))
```

```{r}
occurrence %<>%   
   mutate(decimalLatitude = as.character(format(decimalLatitude, nsmall = 5))) %>%
   mutate(decimalLongitude = as.character(format(decimalLongitude, nsmall = 5))) %>%
   mutate(verbatimLatitude = as.character(format(verbatimLatitude, nsmall = 5))) %>%
   mutate(verbatimLongitude = as.character(format(verbatimLongitude, nsmall = 5)))
```

### continent

```{r}
##occurrence %<>% mutate(continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
#occurrence %<>% mutate(countryCode = "NL") # e.g. "Belgium = BE"
```

### municipality

municipality already in source file

```{r}
#occurrence %<>%
#  mutate(municipality = str_remove(municipality,"\\(BE\\)"))
```

### verbatimLocality

```{r}
occurrence %<>%
  mutate(verbatimLocality = Bereich)


```



### locationID

```{r}
# occurrence %<>% mutate(locationID = SubAreaName)
```

### verbatimCoordinates

```{r}
# occurrence %<>% mutate(locationID = HourSquareName)
```

### verbatimLatitude

### verbatimLongitude

### verbatimCoordinateSystem

```{r}
#occurrence %<>% mutate(verbatimcoordinatesystem = "Amersfoort/RD new") # 
```

### verbatimSRS

```{r}
#occurrence %<>% mutate(verbatimSRS = "EPSG:28992")
```

### decimalLatitude

### decimalLongitude

### geodeticDatum

### coordinateUncertaintyInMeters

### georeferenceRemarks

### identifiedBy

```{r}
occurrence %<>% mutate(identifiedBy = 'ITAW')
```

### scientificName 

```{r}
head(occurrence, n=5)
occurrence %<>% mutate(scientificName = recode(Tier
                   , "Nutria"  = "Myocastor coypus"
                   , "Bisam" = "Ondatra zibethicus"
                   , "Waschbär"  = "Procyon lotor"
                                                    ))
```

### kingdom

```{r}
occurrence %<>% mutate(kingdom = "Animalia")
```

### scientificNameAuthorship

```{r}
occurrence %<>% mutate(scientificNameAuthorship = recode(Tier
                  ,"Bisam" = "Linnaeus, 1766"
                  ,"Nutria"  = "Molina, 1782"
                  ,"Waschbär"  = "Linnaeus, 1758"
                  
                                       ))
```

### verbatimScientificName

```{r}
occurrence %<>% mutate(vernacularName = Tier)
```

# taxonRank

```{r}
occurrence %<>% mutate(taxonRank = "species")
```

# nomenclaturalCode

```{r}
occurrence %<>% mutate(nomenclaturalCode = "ICZN") # e.g. "ICZN"
```

### occurrenceStatus

```{r}
occurrence %<>% 
    ##select(individualCount) %>%
    mutate(occurrenceStatus = case_when(individualCount > 0 ~ "present",
                              individualCount == 0 ~ "absent"
                                 )
                                )
```

## Post-processing

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
occurrence %<>% select(-c(UserName, UserEmail, Datum, Uhrzeit, Tier, "Anmerkung (nicht verpflichtend)",year, month, day, verbatimLongitude, verbatimLatitude, Bereich, Landschaft, Material ))  #remove collection columns
          ##  rename(scientificName = verbatimScientificName) 
```

Define the order of the output columns

```{r}
col_order <- c( "type","language","license","rightsHolder","accessRights","datasetID"
               ,"institutionCode","datasetName","basisOfRecord","occurrenceID","recordedBy"
               ,"individualCount","occurrenceStatus","eventDate","habitat", "continent","countryCode","verbatimLocality"
               
               ,"decimalLatitude","decimalLongitude","geodeticDatum","coordinateUncertaintyInMeters"
               ,"identifiedBy","scientificName","vernacularName","kingdom","scientificNameAuthorship","taxonRank"
               ,"nomenclaturalCode"
               ) 
               # still needed for this dataset 
               # removed for this dataset,"occurrenceRemarks","municipality","samplingProtocol","samplingEffort",                            ,"locationID", ,"verbatimLatitude","verbatimLongitude","verbatimSRS" ,"verbatimCoordinateSystem"

occurrence <- occurrence[, col_order]
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets/invasive-macrofytendata", "data", "processed", "occurrence.csv"), na = "")



```

```{r}
#end
```

