---
title: "Darwin Core Mapping"
subtitle: "Stag Beetle legacy data publication: Stag Beetle"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # for the date
library(ggplot2)        # for map
library(mapview)
```

# Read source data

Create a data frame `input_data1` & `input_data2` from the source data:



```{r}
input_data <- read_csv(here::here("datasets", "stagbeetle-legacy-occurrences", "data", "interim", "Stag-beetle-legacy-refined.csv"))

```



Preview data:

```{r}
input_data %>% head(n = 5)
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% remove_empty("rows")
```

```{r}
input_data %>% head(n = 5)
```


# Occurrence core

```{r}
occurrence <- input_data
```

# Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

# Event

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "complete this")
```
### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "INBO") # e.g. "INBO"
```
### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
#occurrence %<>% mutate(datasetID = "insert doi") 
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "INBO") # e.g. "INBO"
```




### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Legacy observations of the stag beetle (Lucanus cervus) in Belgium") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```


### ownerInstitutionCode

```{r}
occurrence %<>% mutate(dwc_ownerInstitutionCode = "INBO")
```


### eventDate



### eventID


## samplingProtocol



### samplingEffort


# Location

### locationID


```{r}
occurrence %<>%
  mutate(dwc_geodeticDatum = "WGS84") %>%
  mutate(dwc_coordinateUncertaintyInMeters = "30")
  
```


```{r}
occurrence %<>%   
   mutate(dwc_decimalLatitude = as.character(format(dwc_decimalLatitude, nsmall = 5))) %>%
   mutate(dwc_decimalLongitude = as.character(format(dwc_decimalLongitude, nsmall = 5)))
```

### continent

```{r}
occurrence %<>% mutate(dwc_continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode


### municipality

### verbatimLocality

### verbatimcoordinates

### verbatimLatitude

### verbatimLongitude

### verbatimcoordinatesystem


### verbatimSRS

### decimalLatitude

### decimalLongitude

### geodeticDatum

### coordinateUncertaintyInMeters

### georeferenceRemarks

# geologicalContext

## formation

```{r}
occurrence %<>% 
    ##select(individualCount) %>%
    mutate(dwc_occurrenceStatus = case_when(dwc_individualCount > 0 ~ "Present",
                                        dwc_individualCount == 0 ~ "Absent",
                                        is.na(dwc_individualCount) ~ "Absent" ))

```
### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```
### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = "species")
```

### nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN") # e.g. "ICZN"
```

## Post-processing

Only keep the Darwin Core columns:

```{r}
occurrence %<>% select(starts_with("dwc_"))
```

Drop the `dwc_` prefix:

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
```

Preview data:

```{r}
occurrence %>% head()
```

```{r}
occurrence <- mutate_all(occurrence, as.character())
```

```{r}
occurrence %<>% unique()
```

```{r}
occurrence %<>% mutate(ID_row = row_number())
```


Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "stagbeetle-legacy-occurrences", "data", "processed", "occurrence.csv"), na = "")
```





### basisOfRecord



### informationWithHeld

### dataGeneralizations

### occurrenceID

### recordedBy

### individualCount


### occurrenceStatus



## eventID

## identification

## identifiedBy


### CollectionCode


#taxon

### scientificName 



### scientificNameAuthorship








## Post-processing



Visualize data on a map

```{r}
occurrenceMap <- occurrence
```

```{r}
occurrenceMap <- occurrenceMap %>%
  filter(decimalLatitude != " ") %>%           # Eerst spaties verwijderen
  mutate(decimalLatitude = as.numeric(decimalLatitude))  # Dan pas omzetten naar numeriek
```
```{r}
occurrenceMap <- occurrenceMap %>%
  filter(decimalLongitude != " ") %>%           # Eerst spaties verwijderen
  mutate(decimalLongitude = as.numeric(decimalLongitude))  # Dan pas omzetten naar numeriek
```



```{r}
occurrenceMap$decimalLatitude <- as.numeric(occurrenceMap$decimalLatitude)
```

```{r}
occurrenceMap$decimalLongitude <- as.numeric(occurrenceMap$decimalLongitude)
```

```{r}
head(occurrenceMap)
```

```{r}

```



```{r}
head(occurrence)

mapview(occurrenceMap, xcol = "decimalLongitude", ycol = "decimalLatitude", crs = 4269, grid = FALSE)

```


```{r}


head(occurrence)

mapview(occurrenceMap, xcol = "decimalLongitude", ycol = "decimalLatitude", crs = 4269, grid = FALSE)

```

