---
title: "Darwin Core Mapping"
subtitle: "For: Bullfrog occurrences script"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # date transformation
library(googlesheets4)         # read gsheets
library(dplyr)
library(RCurl)         #read csv from github
```

# Read source data

Create a data frame `input_data` from the source data:
The source data was corrected in Excel
Muskrat occurrences opened in openRefine
Obsolete columns removed
some columns renamed to DwC term
File exported to csv

```{r}

x <- getURL("https://raw.githubusercontent.com/inbo/sk-analyse/3_create_datacleaning/interim/afvangsten_per_fuik.csv")
input_interim <- read.csv(text = x)
```


Preview data:

```{r}
input_interim %>% head(n = 5)
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_interim <- input_interim %<>% remove_empty("rows")
```

```{r}
input_interim %>% head(n = 5)
```


# Occurrence core

## Pre-processing

Create a dataframe occurrence data only 

```{r}
occurrence <- input_interim
```

## Remove obsolete columns

```{r}
#occurrence %<>% select(-c(kind_nl,day, month, year))
```

# Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

# Event

### datasetID

```{r}
occurrence %<>% mutate(datasetID = "https://doi.orgXXXXXXX") # e.g. "to complete"
```

### type

```{r}
occurrence %<>% mutate(type = "Event")
```

### language

```{r}
occurrence %<>% mutate(language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(rightsHolder = "ITAW") # e.g. "INBO"
```

### accessRights

```{r}
occurrence %<>% mutate(accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
##occurrence %<>% mutate(datasetID = "https:/to complete") 
```

### institutionCode

```{r}
occurrence %<>% mutate(institutionCode = "ITAW") # e.g. "INBO"
```

### datasetName

```{r}
occurrence %<>% mutate(datasetName = "MICA - Muskrat and Coypu occurrences collected by ITAW in Germany") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

The following terms contain information about the taxon:

### basisOfRecord

```{r}
occurrence %<>% mutate(basisOfRecord = "HumanObservation")
```

### informationWithHeld

### dataGeneralizations

### occurrenceID

**This need to be fixed**

```{r}
occurrence %<>% rename(occurrenceID = 'ID') 
#                mutate(occurrenceID = str_c("UVW:", occurrenceID))
```

### recordedBy

```{r}
occurrence %<>% mutate(recordedBy = 'ITAW')
```

### individualCount

```{r}
occurrence %<>% rename(individualCount = Vangë) 
```

### organismQuantity

### organismQuentityType

### sex
no sex available in this dataset

```{r}
#occurrence %<>% mutate(sex = recode(Gender, ***))
```

### lifeStage

```{r}
#occurrence %<>% mutate(lifeStage = recode(Age,
#"Muskusrat" = "Muskusrat",
#                   "moer"  = "female",
#                   "ram" = "male",
#                   "Onbekend" = "",
#                                   ))
```

### behavior

### occurrenceRemarks

```{r}
# occurrence %<>% rename(occurrenceRemarks = 'action_en')
```

### habitat

```{r}
occurrence %<>% mutate(occurrenceRemarks = recode(Landschaft
                                        
                                      , "Graben"      = "in a hole"
                                      , "Teich/Altarm" = "near Pond / oxbow lake"
                                      , "Bach" = "near stream"
                                      , "Teichanlage (Fischerei)" = "near fishing pond"
                                      , "Binnensee" = "near lake"))
```


### samplingProtocol

```{r}

occurrence %<>% mutate(samplingProtocol = recode(Material,
                    
                            "1" =	 "metal box trap (Grimm)"
                          , "2" =  "shot"
                          , "3" =  "funnel trap (Weka invasiv)"
                          , "4" =  "wire box trap"
                          , "5" =  "Funnel trap (Trapperneozoen)"
                          , "6" =  "wooden box trap"
                          , "7" =  "killing trap (model Hausmann)"
                          , "8" =  "drowning cage trap"
                          , "9" =  "killing trap (model Conibear)"
                          , "10" = "killing trap (model Leprich)"
                          , "11" = "round killing trap"
                          , "12" = "killing trap (model Kerschl)"))

```


### samplingEffort

### eventDate

```{r}
occurrence %<>% mutate(eventDate = ymd(Datum)) 
```

```{r}
occurrence %<>% mutate(year = year(eventDate)) %<>%
                mutate(month = month(eventDate)) %<>%
                mutate(day = day(eventDate)) %<>%
                mutate(eventDate = date(eventDate))
```

# Location

```{r}
head(occurrence)

occurrence %<>% rename(decimalLongitude = X) %>%
                rename(decimalLatitude = Y) %>%
                mutate(geodeticDatum = "WGS84") %>%
                mutate(coordinateUncertaintyInMeters = "30") %>%
                mutate(countryCode = "DE")  %>%            
                mutate(continent = "Europe")
```

```{r}
head(occurrence, n = 5)
occurrence %<>%
#   mutate(verbatimLongitude = round(verbatimLongitude, digits = 0 )) %>%
#   mutate(verbatimLatitude = round(verbatimLatitude, digits = 0)) %>%
   mutate(decimalLongitude = round(decimalLongitude, digits = 5)) %>%
   mutate(decimalLatitude = round(decimalLatitude, digits = 5))
```

```{r}
occurrence %<>%   
   mutate(decimalLatitude = as.character(format(decimalLatitude, nsmall = 5))) %>%
   mutate(decimalLongitude = as.character(format(decimalLongitude, nsmall = 5))) %>%
   mutate(verbatimLatitude = as.character(format(verbatimLatitude, nsmall = 5))) %>%
   mutate(verbatimLongitude = as.character(format(verbatimLongitude, nsmall = 5)))
```

### continent

```{r}
##occurrence %<>% mutate(continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
#occurrence %<>% mutate(countryCode = "NL") # e.g. "Belgium = BE"
```

### municipality

municipality already in source file

```{r}
#occurrence %<>%
#  mutate(municipality = str_remove(municipality,"\\(BE\\)"))
```

### verbatimLocality

```{r}
occurrence %<>%
  mutate(verbatimLocality = Bereich)


```



### locationID

```{r}
# occurrence %<>% mutate(locationID = SubAreaName)
```

### verbatimCoordinates

```{r}
# occurrence %<>% mutate(locationID = HourSquareName)
```

### verbatimLatitude

### verbatimLongitude

### verbatimCoordinateSystem

```{r}
#occurrence %<>% mutate(verbatimcoordinatesystem = "Amersfoort/RD new") # 
```

### verbatimSRS

```{r}
#occurrence %<>% mutate(verbatimSRS = "EPSG:28992")
```

### decimalLatitude

### decimalLongitude

### geodeticDatum

### coordinateUncertaintyInMeters

### georeferenceRemarks

### identifiedBy

```{r}
occurrence %<>% mutate(identifiedBy = 'ITAW')
```

### scientificName 

```{r}
head(occurrence, n=5)
occurrence %<>% mutate(scientificName = recode(Tier
                   , "Nutria"  = "Myocastor coypus"
                   , "Bisam" = "Ondatra zibethicus"
                   , "Waschbär"  = "Procyon lotor"
                                                    ))
```

### kingdom

```{r}
occurrence %<>% mutate(kingdom = "Animalia")
```

### scientificNameAuthorship

```{r}
occurrence %<>% mutate(scientificNameAuthorship = recode(Tier
                  ,"Bisam" = "Linnaeus, 1766"
                  ,"Nutria"  = "Molina, 1782"
                  ,"Waschbär"  = "Linnaeus, 1758"
                  
                                       ))
```

### verbatimScientificName

```{r}
occurrence %<>% mutate(vernacularName = Tier)
```

# taxonRank

```{r}
occurrence %<>% mutate(taxonRank = "species")
```

# nomenclaturalCode

```{r}
occurrence %<>% mutate(nomenclaturalCode = "ICZN") # e.g. "ICZN"
```

### occurrenceStatus

```{r}
occurrence %<>% 
    ##select(individualCount) %>%
    mutate(occurrenceStatus = case_when(individualCount > 0 ~ "present",
                              individualCount == 0 ~ "absent"
                                 )
                                )
```

## Post-processing

filter for 0 lat-long

```{r}
head(occurrence)

occurrence %<>% filter(decimalLongitude != " 0.00000")
```


```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
occurrence %<>% select(-c(UserName, UserEmail, Datum, Uhrzeit, Tier, "Anmerkung (nicht verpflichtend)",year, month, day, verbatimLongitude, verbatimLatitude, Bereich, Landschaft, Material ))  #remove collection columns
          ##  rename(scientificName = verbatimScientificName) 
```

Define the order of the output columns

```{r}
col_order <- c( "type","language","license","rightsHolder","accessRights","datasetID"
               ,"institutionCode","datasetName","basisOfRecord","samplingProtocol", "occurrenceID","recordedBy"
               ,"individualCount","occurrenceStatus","occurrenceRemarks","eventDate","continent","countryCode","verbatimLocality"
               
               ,"decimalLatitude","decimalLongitude","geodeticDatum","coordinateUncertaintyInMeters"
               ,"identifiedBy","scientificName","vernacularName","kingdom","scientificNameAuthorship","taxonRank"
               ,"nomenclaturalCode"
               ) 
               # still needed for this dataset 
               # removed for this dataset,"occurrenceRemarks","municipality","samplingProtocol","samplingEffort",                            ,"locationID", ,"verbatimLatitude","verbatimLongitude","verbatimSRS" ,"verbatimCoordinateSystem"

occurrence <- occurrence[, col_order]
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets/mica-itaw-occurrences", "data", "processed", "occurrence.csv"), na = "")



```

```{r}
#end
```

